### 2015-07-09 narrative/qos.tex
tied together and smoothed wording a bit.

\subsubsection{Resource Management}
There is an important connection between finding the resources
that can fulfill a particular request and the performance predictability of that request. As such, this demands
careful resource management strategies that can deal with placement and retrieval of 
a large number of data chunks spread across the entire storage hierarchy. 
%This connection is particularly important for data retrieval 
%- in large scale storage systems, finding a resource to read can be a significant, if not dominant, 
%component of overall read latency. 
Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
scalability, because they require a single entity that tracks updates in the system. 
Traditional file systems are good examples for this ; most updates to files require updates 
to their centrally maintained block allocation tables. 
The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, searchs for 
resources at request time. Pile systems allow very efficient writes because clients can write wherever
there is free space and there is very low bookkeeping overhead. However, unbounded search in pile systems
is possible, e.g., when it is unclear whether the resource actually exists. 
Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 
partitioning of the search space and a method to limit the search to a small number of partitions.

Frequently this involves a hash function that is relatively stable against changes to its range, 
such as failures or updates to the hardware of the system. Consistent hashing~\cite{karger:stoc97}
is scalable since it can be well-known throughout a system of any size and only requires minimal 
movement of data in the presence of failures or system changes. 
%Consistent hash functions can also
%have sets of ordered but infinite lists as their range (which works well with primary replication)
%~\cite{honicky:ipdps04}, implement declustering (limiting overlap between partitions), and support
%placement policy rules~\cite{weil:sc06}.

\paragraph{Approach:} 
In this proposal we will investigate the trade-off in terms of predictable performance between 
``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' 
data, pushing the cost of finding to the moment when the data is actually needed. This trade-off 
has a parallel in data management where there are two alternative paths to achieve knowledge: 
the analogue to filing is  to ``extract, transform, and load'' (ETL) data into a warehouse, paying 
an upfront ingest cost to achieve fast querying; and the analogue to piling is to not pay any upfront 
cost and later querying   the raw data directly by using a data processing platform such as Hadoop.  Recent work 
has found that the combination of these two techniques can be optimized to perform better than 
any one technique alone~\cite{lefevre:sigmod14a}. We believe that a similar opportunity exists 
for combining filing and piling approaches to achieve performance predictability without sacrificing 
too much flexibility.

More specifically, the coarseness of a search space partitioning limits the search time within 
a partition, i.e. finely partitioned search spaces require little or no search within a partition 
but offer little placement flexibility. Partitioning can also be fine in one dimension, and coarse 
in another to allow flexibility where it is most needed while still ensuring fast finding. For 
example, a hash function can finely partition a low storage tier but each partition can extend 
across multiple tiers so that a search still has to occur in order to find a resource on a particular 
tier.

@@

\subsubsection{Quality of Service (QoS)}
QoS refers to the
properties of the performance quality of a particular service, in
this case the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS \emph{reservations}, rather than continuing the current model
of allowing applications to compete freely without restriction, for part or all of the storage
resources. This traditional approach leads to interference effects
\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact I/O performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
Unfettered, this competition can yield a lower system and application performance 
than the case where applications attempt to reserve storage resources, thereby enabling the storage system
to reason over its performance in terms of reservations and actual
performance.
In the cases where an application arrives that needs to run more urgently
than any of the other already running applications, the system
can rapidly respond to it by reducing the amount of resources taken 
by less urgent applications, allowing the urgent application to finish,
and later reverting back to the previous committed resource level for the rest.
This can be automatically enforced via user- (or administrator-) specified \textit{policies}\S\ref{sec:data-description}.

There are two fundamental causes that can impede the I/O performance
in an extreme-scale SSIO system. One is indirection, where layers
of storage tiers are employed either for performance and/or for
scalability reasons.  The direct consequence is that there could
be multiple traversing paths from application end to the rest place
of data, and more often than not, the I/O paths are not under control
under any single authority. The other cause is the shared use of
resources. The best effort I/O request/response nature and lack of
QoS mechanisms imply that there is little guarantee in terms of
expected performance. Both indirection and shared use of resource
contribute to a high probability of imbalanced use of resources
thus the occurrence of congestion and degraded performance.

As an example to demonstrate that this disconnection will hurt
system performance: We launch 4096 processes with each process doing
a single file I/O operation against half of the Spider II file
system. The traces of those files are analyzed to examine the
utilization distribution of different components. Figure (a), (b)
and (c) shows the resource usage distribution for OSTs, OSSes, and
LNETs, respectively. We observe that there exists a significant
variation in usage across components of any given type (e.g., OST,
OSS or LNET). For example, some OSTs are used more than 10 times
while some others are never used (corresponding to zero frequency
count). Similarly, OSSes and LNETs show significant imbalance in
usage under the default placement strategy. Consequently, imbalanced
resource utilization increases the contention at certain components
more than others.

@@

% This is about finding the relationshiop of workload-dependent to workload-independent metrics. -- Carlos
 %are driven from multiple different approaches. First, supporting
%capability runs is a priority for capability machines. We propose to address
%capability run performance by incorporating monitoring for prepartory runs at
%smaller scale to profile the application output characteristics both from a
%writing velocity and volume, but also for example read patterns for the data
%analytics required to generate scientific insights from the raw data.

@@

\paragraph{Approach:} The goals of this proposal require applications
to demand guarantees in terms of absolute performance qualities.
The first key enabler for any service to make absolute guarantees
is \emph{performance isolation}, i.e. the ability to shield the
performance impact of one task from another. The most straight-forward
but least flexible strategy is to simply cap the latency of each
task, regardless of its completeness (see for
example~\cite{decandia:sosp07}). This strategy is applicable when
incomplete tasks still have value and missing results are either
not important or can be easily retrieved in subsequent tasks. A
more generally applicable strategy is to control the admission of
tasks to the system: \emph{admission control} maintains a
\emph{utilization model} that can predict the utilization of system
resources given a new task, estimates the current utilization of
the system, and decides whether the task can be admitted without
overloading the system. Thus, admission control avoids system
overload conditions that lead to hard-to-model chaotic behaviors.
Consequently, utilization models can frequently be approximated by
automatically calibrated linear models (see for
example~\cite{skourtis:hpdc12}).

Another essential component of performance isolation is the
\emph{charging model} which determines which task(s) should be
charged how much for a request. Performance isolation is implemented
by the charging model by ensuring that performance of a task only
depends on its reservation and its workload behavior (but not on
other workloads in the system). Each task has a budget based on its
reservation and spends it based on the cost of each request determined
by the charging model. A charging model determines the cost based
on the utilization model and the interactions a request can have
with other requests (e.g. a random request to spinning media
interrupting a stream of sequential requests).

Performance isolation (including its utilization and charging models)
critically depends on \emph{workload-independent metrics}. Even
though throughput and latency are frequently the most meaningful
metrics for an application, they are not workload-independent and
therefore are inadequate for measuring the utilization of a device
across all workloads, unless one accepts worst-case estimates that
can lead to gross underutilization of resources (e.g. the throughput
of random and sequential workloads in spinning media differ by
orders of magnitudes). A common workload-independent metric is
\emph{time utilization}, i.e. the amount of time a resource is
utilized in a given time interval. Time utilization has the advantage
of having an always defined maximum and therefore makes resources
fully reservable. Even though a workload-independent metric might
at first not appear meaningful for an application, given a particular
workload an application can discover the relationship between its
workload-dependent metric and the workload-independent metric.
Because of performance isolation, the application has to discover
this relationship only once.
To achieve effective performance isolation, we seek to expand the resource management
capabilities in Sirocco to support additional utilization 
metrics to guide resource management
decisions. Currently, Sirocco offers automatic resource management capabilities,
based on resource capacity and data resilience requirements, and it needs to be extended
to offer data-centric metrics that help guide the forced QoS decisions.

@@

\paragraph{Related Work:} On current production HPC systems, absolute I/O performance
guarantees are not possible. Previous work \cite{lofstead:2010:io-variability,liu_hotstorage} 
from this project team uses a client-assisted approach to mitigate I/O variablity at scale (100,000-core).
Recently server side QoS scheduling has also been studied for HPC applications \cite{Dai:2014}.
However, the key hurdle of scalability has not been solved.
Existing efforts on scheduing resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not
require modification to use new storage access APIs. By using a priority
tagging system, we will offer defaults suitable for all applications that can
be informed by simple extensions to the job scheduler or more advanced
management through new APIs. On the other hand, storage QoS has been
studied for enterprise applications and systems \cite {Gulati:2007,Gulati:2010,Gulati:2012}
in virtualized environments. It is not straightforward to adapt these solutions
to HPC storage systems because of the scalability and efficiency concerns, despite
they achieved very good performance isolations.

>>>>

\subsubsection{Resource Management}
Careful resource management strategies that can deal with placement and
retrieval of a large number of data chunks spread across the entire storage
hierarchy are key for predictable performance. 
%This connection is particularly important for data retrieval - in large scale
%storage systems, finding a resource to read can be a significant, if not
%dominant, component of overall read latency. 
Locating resources is typical a distributed systems problem. Centralized
solutions suffer scalability challenges due to centralized, serialized system
updates.  POSIX-compliant file systems are good examples of this
strucutre--most updates to files require updates to their centrally maintained
block allocation tables.

An opposite example of this central management strategy are ``pile'' systems,
which, instead of tracking updates, they search for resources upon request.
Pile systems feature efficient writes because clients can write wherever there
is free space with low bookkeeping overhead.  However, unbounded search is
possible, e.g., when it is unclear whether the resource actually exists.
Methods to make searches more efficient without reverting to centralized
bookkeeping usually involves search space partitioning and a method to limit
the search to a small number of partitions.  Frequently this involves a
hash function that is relatively stable against changes to its range, such as
failures or updates to the hardware of the system. Consistent
hashing~\cite{karger:stoc97} is scalable since it can be well-known throughout
a system of any size and only requires minimal movement of data in the presence
of failures or system changes. 
%Consistent hash functions can also have sets of ordered but infinite lists as
%their range (which works well with primary replication)
%~\cite{honicky:ipdps04}, implement declustering (limiting overlap between
%partitions), and support placement policy rules~\cite{weil:sc06}.

Sirocco offers more of a pile struture rather than a file structure. One of the
challenges we face is offering performance predictability more similar to
POSIX-compliant file systems while we take advantage of the distributed,
flexible storage management offered by Sirocco.

\paragraph{Approach:} 
In this proposal, we will investigate the trade-off in terms of predictable
performance between ``filing'' and ``piling'' data.  This trade-off has a
parallel in data management where there are two alternative paths to achieve
knowledge. The analogue to filing is  to ``extract, transform, and load'' (ETL)
data into a warehouse paying an upfront ingest cost to achieve fast querying.
The analogue to piling is to not pay any upfront cost and later querying the
raw data directly by using a data processing platform such as Hadoop.  Recent
work has found that the combination of these two techniques can be optimized to
perform better than any one technique alone~\cite{lefevre:sigmod14a}. We
believe that a similar opportunity exists for combining filing and piling
approaches to achieve performance predictability without sacrificing too much
flexibility. Our additional challenge is incorporating both the storage tiers
and potential data quality metrics.

More specifically, the coarseness of a search space partitioning limits the
search time within a partition, i.e., finely partitioned search spaces require
little or no search within a partition but offer little placement flexibility.
Partitioning can also be fine in one dimension and coarse in another to allow
flexibility where it is most needed while still ensuring fast finding. For
example, a hash function can finely partition a low storage tier but each
partition can extend across multiple tiers so that a search must occur in order
to find a resource on a particular tier.

@@

\subsubsection{Quality of Service (QoS)}
Our ultimate goal, rather than just performance predictability, is to offer
predictable Quality of Service (QoS) for the storage system.
%Performance quality can be expressed in either relative or absolute
%terms. Examples of relative performance terms are ``fair'',
%``proportional'', or ``priority/class-based'' (``the higher the
%priority the better the service'' or ``1st class is better than
%2nd''). The key advantage of relative performance terms is that
%they are easy to implement. The key disadvantage is the difficulty
%of making any guarantees based on those implementations --
%even relative guarantees are mired by the well-known effect of
%priority inversion~\cite{lampson:cacm80}. Examples of absolute
%performance terms are ``rate-based'', ``soft real-time'', and
%``hard real-time''. The key advantage of absolute performance
%terms are that they enable the implementation of strong
%guarantees: absolute terms enable agreements between application
%tasks and storage that we call \emph{reservations} and that do not
%change meaning depending
%on the storage system's workload. The key challenge of absolute
%performance terms is that their implementation is non-trivial,
%especially in large-scale systems.
One of the key differentiating features this project offers is driving storage
decisions through QoS~\emph{reservations} rather than continuing the current
model of allowing applications to compete freely without restriction for part
or all of the storage resources. The current approach leads to interference
effects~\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly
impact I/O performance predictability. 
%Further, with the introduction of additional tiers in the
%storage stack, the performance variability will increase with greater
%competiion for limited high performance resources. 
%% Temporaily commented out the above sentence, because it may not be true for
%the next DOE systems. 
Unfettered, this competition can yield lower system and application performance
than if applications attempt to reserve storage resources thereby enabling the
storage system to reason over its performance in terms of reservations and
actual performance.  In the cases where an application arrives that needs to
run more urgently than any of the other already running applications, the
system can rapidly respond to it by reducing the amount of resources taken by
less urgent applications, allowing the urgent application to finish, and later
reverting back to the previous committed resource level for the rest.  This can
be automatically enforced via user- or administrator- specified
\textit{policies}\S\ref{sec:data-description}.

There are two fundamental causes that can impede the I/O performance in an
extreme-scale SSIO system. One is indirection, where layers of storage tiers
are employed either for performance and/or for scalability reasons.  The direct
consequence is that there could be multiple traversing paths from application
end to the data resting place, and more often than not, the I/O paths are not
under control under of any single authority. The other cause is shared
resources use. The best effort I/O request/response nature and lack of QoS
mechanisms imply that there is little expected performance guarantee.  Both
indirection and shared resources use contribute to a high probability of
imbalanced resources use and thus the occurrence of congestion and degraded
performance.

Consider this example demonstrating this disconnection will hurt system
performance. We launch 4096 processes with each process doing a single file I/O
operation against half of the Spider II file system at ORNL. The file traces
are analyzed to examine the utilization distribution of different components.
Figures (a), (b) and (c) show the resource usage distribution for OSTs, OSSes,
and LNETs, respectively. We observe a significant usage variation
across components of any given type (e.g., OST, OSS or LNET). For example, some
OSTs are used more than 10 times while some others are never used
(corresponding to zero frequency count). Similarly, OSSes and LNETs show
significant usage imbalance under the default placement strategy.
Consequently, imbalanced resource utilization increases the contention at
certain components more than others.

@@

% This is about finding the relationshiop of workload-dependent to
% workload-independent metrics. -- Carlos
%are driven from multiple different approaches. First, supporting
%capability runs is a priority for capability machines. We propose to address
%capability run performance by incorporating monitoring for prepartory runs at
%smaller scale to profile the application output characteristics both from a
%writing velocity and volume, but also for example read patterns for the data
%analytics required to generate scientific insights from the raw data.

@@

\paragraph{Approach:}
The goals of this proposal require applications to demand guarantees in terms
of absolute performance qualities.  The first key enabler for any service to
make absolute guarantees is \emph{performance isolation}, i.e., the ability to
shield the performance impact of one task from another. The most
straight-forward but least flexible strategy is to simply cap the latency of
each task, regardless of its completeness (see for
example~\cite{decandia:sosp07}). This strategy is applicable when incomplete
tasks still have value and missing results are either not important or can be
easily retrieved in subsequent tasks. A more generally applicable strategy is
to control task admission to the system. \emph{Admission control} maintains a
\emph{utilization model} that can predict the utilization of system
resources given a new task, estimates the current utilization of the system,
and decides whether the task can be admitted without overloading the system.
Thus, admission control avoids system overload conditions that lead to
hard-to-model chaotic behaviors.  Consequently, utilization models can
frequently be approximated by automatically calibrated linear models (see for
example~\cite{skourtis:hpdc12}).

Another essential component of performance isolation is the \emph{charging
model} that determines which task(s) should be charged how much for a request.
Performance isolation is implemented by the charging model by ensuring that
task performance only depends on its reservation and its workload behavior (but
not on other workloads in the system). Each task has a budget based on its
reservation and spends it based on the cost of each request determined by the
charging model. A charging model determines the cost based on the utilization
model and the interactions a request can have with other requests (e.g. a
random request to spinning media interrupting a stream of sequential requests).

Performance isolation, including its utilization and charging models,
critically depends on \emph{workload-independent metrics}. Even though
throughput and latency are frequently the most meaningful metrics for an
application, they are not workload-independent and therefore are inadequate for
measuring device utilization across all workloads unless one accepts
worst-case estimates that can lead to gross resource underutilization. For
example, the throughput of random and sequential workloads in spinning media
differ by orders of magnitudes. A common workload-independent metric is
\emph{time utilization}, i.e., the amount of time a resource is utilized in a
given time interval. Time utilization has the advantage of having an always
defined maximum and therefore makes resources fully reservable. Even though a
workload-independent metric might at first not appear meaningful for an
application, given a particular workload an application can discover the
relationship between its workload-dependent metric and the workload-independent
metric.  Because of performance isolation, the application has to discover this
relationship only once.  To achieve effective performance isolation, we seek to
expand the resource management capabilities in Sirocco to support additional
utilization metrics to guide resource management decisions. Currently, Sirocco
offers automatic resource management capabilities, based on resource capacity
and data resilience requirements, and it needs to be extended to offer
data-centric metrics that help guide the forced QoS decisions.

@@

\paragraph{Related Work:}
On current production HPC systems, absolute I/O performance guarantees are not
possible. Previous work~\cite{lofstead:2010:io-variability,liu_hotstorage} from
this project team uses a client-assisted approach to mitigate I/O variablity at
scale (100,000-core).  Recently, server side QoS scheduling has also been
studied for HPC applications~\cite{Dai:2014}.  However, the key hurdle of
scalability has not been solved.  Existing efforts on scheduing
resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not require
modification to use new storage access APIs. By using a priority tagging
system, we will offer defaults suitable for all applications that can be
informed by simple extensions to the job scheduler or more advanced management
through new APIs. On the other hand, storage QoS has been studied for
enterprise applications and systems~\cite{Gulati:2007,Gulati:2010,Gulati:2012}
in virtualized environments. It is not straightforward to adapt these solutions
to HPC storage systems because of the scalability and efficiency concerns,
despite they achieved very good performance isolations.
